### 2020-05-29 text/si_vedrovice.Rmd
text and SI - methods

---
title           : "Supplementary Information: Vedrovice burial analysis"
author          : "Petr Pajdla"
    # name          : "Petr Pajdla"
    # affiliation   : "Department of Archaeology and Museology,
    #                  Masaryk University, Brno, Czech Republic"
    # email         : "petr.pajdla@gmail.com"
date            : "`r format(Sys.time(), '%d %B, %Y')`"
output:
    bookdown::pdf_document2:
        latex: pdflatex
        fig_caption: true
        number_sections: false
        keep_tex: false
bibliography    : "bib_vedrovice.bib"
csl             : "european-journal-of-archaeology.csl"
---

# Randomness of Artefact Combinations

For pair of artefacts $i$ and $j$, the observed and expected number of co-occurences are
denoted $o_{ij}$ and $e_{ij}$ respectively.
$R$ is the number of variables.
The $S$ statistic, the overall deviation from expected number of co-occurences for
all artefacts is defined as the mean square of deviations:

\begin{equation}
S = \sum_{i=1}^{R}\sum_{j=1}^{R}\frac{(o_{ij} - e_{ij})^2}{R^2} (\#eq:sstat)
\end{equation}

also:\@ref(eq:sstat)

$$ S = \sum_{i=1}^{R} \frac{v_{i}}{R}$$

where

$$ v_i = \sum_{j=1}^{R} \frac{(o_{ij} - e_{ij})^2}{R}$$

## Implementation