### 2020-08-18 text/si_vedrovice.Rmd
improve figures and tables

---
title           : "Supplementary Information: Vedrovice burial analysis"
author          : "Petr Pajdla"
    # name          : "Petr Pajdla"
    # affiliation   : "Department of Archaeology and Museology,
    #                  Masaryk University, Brno, Czech Republic"
    # email         : "petr.pajdla@gmail.com"
date            : "`r format(Sys.time(), '%d %B, %Y')`"
output:
    bookdown::pdf_document2:
        latex: pdflatex
        fig_caption: true
        number_sections: false
        keep_tex: false
bibliography    : "bib_vedrovice.bib"
csl             : "european-journal-of-archaeology.csl"
---

```{r}
library(here)
library(tidyverse)
library(knitr)
library(kableExtra)
```

# Spatial analysis

<!--
If $s_i$ is an event location, then
$G$ function is given by the formula
\@ref(eq:gfun)

\begin{equation}
\label{eq:gfun}
G(d) = \frac{\#[d_{min}(s_i)<d]}{n}
\end{equation}

$F$ function by
$$F(d) = \frac{\#[d_{min}(p_i, S)<d]}{m}$$
where ${p_1 \cdots p_i \cdots p_m}$ is a set of $m$ randomly selected locations.
Ripley's $K$ function
$$K(d) = \frac{\alpha}{n}\frac{1}{n}\sum_{i=1}^n \#[S \in C(s_i, d)]$$
$T$ function
$$ $$
-->

$s_i$ is an event location.

## G function

$$ G(d) = \frac{\sum[d_{min}(s_i)<d]}{n} $$

## F function

$$F(d) = \frac{\#[d_{min}(p_i, S)<d]}{m}$$

where ${p_1 \cdots p_i \cdots p_m}$ is a set of $m$ randomly selected locations.

## Ripley's K function

$$K(d) = \frac{\alpha}{n}\frac{1}{n}\sum_{i=1}^n \#[S \in C(s_i, d)]$$

# Randomness of Artefact Combinations

For pair of artefacts $i$ and $j$, the observed and expected number of co-occurences are
denoted $o_{ij}$ and $e_{ij}$ respectively.
$R$ is the number of variables.
The $S$ statistic, the overall deviation from expected number of co-occurences for
all artefacts is defined as the mean square of deviations:

\begin{equation}
S = \sum_{i=1}^{R}\sum_{j=1}^{R}\frac{(o_{ij} - e_{ij})^2}{R^2} (\#eq:sstat)
\end{equation}

also:\@ref(eq:sstat)

$$ S = \sum_{i=1}^{R} \frac{v_{i}}{R}$$

where

$$ v_i = \sum_{j=1}^{R} \frac{(o_{ij} - e_{ij})^2}{R}$$

## Implementation

# Results

```{r, echo = FALSE, message = FALSE, results = 'asis'}
tab <- read_csv(here("data/temp/nb_sex_p.csv")) %>% 
         mutate(relation = str_c(from, to, sep = " - ")) %>% 
  select(relation, everything(), -from, -to)

tnames <- c("relation", rep(c("mean", "p"), 3))

kable(tab, booktabs = TRUE, col.names = tnames) %>% 
  add_header_above(c(" ", "Buffer" = 2, "Delaunay" = 2, "Gabriel" = 2))
```