### 2020-09-25 text/article_vedrovice.Rmd
finish spatial analysis

---
title           : "Spatial Analysis and Grave Goods Distribution at the Cemetery of Vedrovice"
# "A resampling approach to spatial arrangement and contents of burials in the Vedrovice cemetery"

author          : "Petr Pajdla"
    # name          : "Petr Pajdla"
    # affiliation   : "Department of Archaeology and Museology,
    #                  Masaryk University, Brno, Czech Republic"
    # email         : "petr.pajdla@gmail.com"
date            : "`r format(Sys.time(), '%d %B, %Y')`"
abstract        :
  "Spatial patterns in cemeteries and grave goods associations at the
  cemetery of Vedrovice - Široká u Lesa are explored using computationaly
  instensive formal methods.
  Resampling approaches both to intrasite spatial arrangement and
  distribution of grave goods are explored in
  relation to layers of identity represented by sex, age and origin of the
  buried bodies.
  As a result of the investigation we conclude that...  "

output:
    bookdown::pdf_document2:
    # bookdown::word_document2:
        latex: pdflatex
        fig_caption: true
        number_sections: false
        keep_tex: false
bibliography    : "bib_vedrovice.bib"
csl             : "european-journal-of-archaeology.csl"
---

<!-- note: length - approx. 10 - 12 pages (no bibliography, tables, figures) -->
<!-- i.e. 18 000 - 22 000 characters -->

```{r init, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, message = FALSE, warning = FALSE)
# library(tidyverse)
library(here)
```

<!-- ======================================================================= -->

# Introduction

The Cemetery of Vedrovice is one of the earliest and most intensively studied
*Linearbandkeramik* (LBK) cemeteries.
Recent contributions use the evidence from the cemetery in a broader discussion
of post-marital residence patterns [@hrncir2020] and
sexual inequalities and gender in the Neolithic [@masclanslatorre2020; @bickle2020].
Here we focus on exploration of relationship between three proposed *layers* of
identity, biological sex and age of the body determined from sekeletal remains
by anthropological methods and origin of the body based on Strontium isotope
signal.
We explore whether the structure observed in these three categories is
mirrored in the grave goods associations and spatial organization within the
cemetery.
The analysis is set into a broader discussion of gender relations and identity
inside Neolithic communities.

<!-- view of gender in the Neolithic, archaeological context -->
@robb2017 [128] noted lack of gender archaeology debate for the
European Neolithic in comparison with other periods.
Authors present gender in the Neolihic as being
of a very different nature than gender in the later prehistory.
Gender in the Neolithic is not clearly expressed and
biological differences between sexes are not transformed into binary cultural
categories (ibid. 140-142).
Considering the Neolithic burial evidence across Europe, authors find that it
'shows very little gender distinction; when it does, it is rarely a clear-cut
binary categorization but a matter of degree and overlap, sometimes polythetically
combined with factors such as age and localness' (ibid. 138).
In response, @bickle2020 [14] argues that 'biological sex may not have defined gender
during the LBK' although the 'sexed bodies did matter'.

<!-- We focus on relations between burials and bodies in terms of biological
sex, age and origin, exceptionality of grave goods and location within the
cemetery.-->
We focus predominantly on investigating non-random patterns in artefact
co-occurences and spatial arragement of the cemetery in relation to the
proposed *layers* of Neolithic identity, i.e. sex, age and localness of the body.
Our question is thus how are grave good assemblages related to aforementioned
categories and whether there is a consistent relationship between
these categories forming the *identity*, grave goods distributions and spatial
arrangement of the cemetery.

<!-- gender relations in LBK -->
<!-- Although gender relations in the LBK might not be -->

<!-- resampling and computer-intensive techniques-->
The approach taken is a quantitative one embracing resampling techniques and
computer-intensive methods in general.
The resampling techniques were introduced to mortuary studies by
@manly1996 but the application is still limited [@sosna2008, 343].
As noted by Sosna (ibid.), the most important benefit of resampling is in its
ability to distinguish random and non-random patterns.
This is crucial for mortuary analysis because the samples studied, including
the one presented here, are usually small in size and identifying which
observed patterns are caused by mere chance associated with small numbers
and which are actually present.

<!-- ### Questions
* Is there evidence that certain artefacts covary,
    either positively or negatively?
    Do certain artefacts tend to occur together, or are certain combinations
    absent or rare?
* How can the presence and absence data best display which burials
    contain similar artefacts,
    and which artefacts have similar distributions over burials?
* *How can the data be analysed so as to show most clearly the relationships*
    *between the different* **burials** in terms of their **contents**?
* *How can the data be analysed so as to show most clearly the relationship*
    *between the different* **artefacts** in terms of their
    **distributions in graves**?
* How are grave good assemblages related to age and sex?
(... variations in burial rite coinciding with age or sex of the deceased ...) -->

<!-- ======================================================================= -->

# Material and methods

<!-- site -->
Vedrovice site is situated in the south of Moravia (Czech Republic).
Earliest known finds and archaeological explorations here are from
the 19th century but the excavation took place in the second half of 20th century.
There are two known burial areas, *Za Dvorem* and larger one *Široká u Lesa* and
a settlement area.
For more details on the site refer to @masclanslatorre2020 or the
original publication [@podborsky2002].
The material analyzed here comes from Vedrovice *Široká u lesa* cemetery.

<!-- sample (burials, artefacts) -->
Originally, this burial ground consisted of at least 108 burials.
Some of the burials were disturbed by later settlement and burial pits in the
area of the cemetery, amateur excavations during the 19th century or later
agricultural activities [@podborsky2002].
These burials are excluded from most of the analysis due to missing and/or
incomplete data and only a sample of 78 well-preserved LBK burials is analyzed here.
Most of the data was obtained from the original publication [@podborsky2002],
anthropological determination of sex and age of the bodies were corrected based on
analysis by @dockalova2008. Carbon, nitrogen and strontium isotope
ratios were collected in @whittle2013 and @richards2008.
Further details on use and function of the artefacts were gathered from
@masclanslatorre2020.
Information on localness of different bodies was adapted from @hrncir2020.

Each burial was originally characterized by 20 variables recording counts of
artefacts present in the graves.
Some of  these original variables are omitted in the analysis due to their rare occurrence,
for instance spondylus arm ring is present in a single burial only,
thus being of no use in establishing relationships between pairs of burials.
Only grave goods occuring in more than 5% of the graves are used in the analysis,
this step reduces the count of the variables to 14.
These include variables noting quantities of various types of pottery vessels in
the graves, counts of chipped stone tools made of either local or non-local raw materials,
presence of polished and beveled stone adzes, grinding slabs, pebbles,
bone tools and different kinds of personal ornament.
The personal ornament includes beads made of spondylus shell,
spondylus pendants or oval shaped buckles.
Last but not least presence of ochre pigment in the grave is also noted.
Quantities of various *grave goods* are simplified into a binary matrix indicating
presence or absence of a given trait.

Another category of variables is derived from retrieved human skeletal remains.
These include biological sex, age simplified into age groups and indicator
variable whether the Strontium signal in the remains is of local origin.
The variable representing the recorded sex is simplified into the following
categories: female-sexed body (abbreviated *F*), male-sexed body (*M*),
non adult (*n. a.*) and indetermined sex (*ind.*).
Age of the body is also simplified into three categories: juveniles (*juv.*)
comprising of groups Infans and Juvenis, adults (*ad.*, all bodies aged as Adultus)
and mature bodies (*mat.*, anthropological categories Adultus-Maturus, Maturus and
Senilis) and bodies of indetermined age (*ind.*).
Last category is the indicator for the origin of the buried person.
Local Strontium range is defined as two standard deviations from mean of Sr ranges of
bodies with age category determined as Infans.
Bodies faling within this range are labeled as local, outside the range as non-local
and where data on Sr values are missing origin is not determined (*ind.*).

<!-- methods -->

## Randomness of artefact combinations

<!--
This study takes advantage of computer-intensive resam-
pling techniques that are based on simulations of the behavior
of statistical models through the generation of random samples
(Baxter, 2003; Efron and Tibshirani, 1993; Manly, 1991). The
strength of these techniques lies in the potential to effectively
distinguish between random and non-random patterns and to
test differences between samples without making assumptions
about the population from which these samples were drawn
(Drennan and Peterson, 2004). These techniques proved to
be very effective for the analysis of patterns in cemeteries
(Manly, 1996).
@sosna2008 -->

To determine non-randomness of various artefact combinations in burials we conduct a
permutation test on the presence/absence data matrix.
The method based on generalized Monte Carlo tests was
introduced to archaeology by @manly1996 and adapted by @sosna2008.
Our approach is based on both of these works.
Original co-occurence matrix is permutated *n = 9999* times with constant
row and column totals, i.e. both the count of artifacts in individual graves
and sum of the given type of artifact across the burials are fixed.

To examine which associations of variables, i.e. artefacts, are infrequent or even not present
and which tend to appear together more often than it would be expected by random chance
the co-occurences of observed and simulated data
are summarized using two statistics suggested by @manly1996.
The *S* statistic is a measure of overall deviation between expected and observed
number of artefact occurences while
the *v* statistic is a measure of difference between co-occurences of
individual artefact with other artifacts
[@manly1996, 475, for details see SI].
To assess relations between groups based on sex, age and origin
statistics introduced by @sosna2008 are used.

<!-- ## Exceptionality index -->

<!-- Exceptionality index (*EI*) is used in order to induce relative exceptionality -->
<!-- of individual burials based on their attributes. -->
<!-- This approach is independent on a prior selection of attributes of importance -->
<!-- or *prestige*. -->
<!-- It was defined by @demjan2015[367] based on a Root Mean Square (*quadratic mean*) of -->
<!-- PCA factor scores: -->
<!-- $$ EI_{attr_g}=\sqrt{\frac{1}{n}\sum_{i=1}^{n}s_i^2} $$ -->
<!-- where $s_i$ are factor scores of individual graves. The mean value is -->
<!-- normalized to range $[0-1]$: -->

<!-- $$ EI_{attr_g}=\frac{EI_{attr_g}}{max~EI_{attr_g}} $$ -->

<!-- spatial analysis -->
## Spatial analysis

Spatial patterning in prehistoric burial grounds is often examined to find
any internal structure, arrangement or possible grouping.
We approach the task as a point pattern analysis problem given that
a burial (event) is simplified as its centroid location in a two dimensional space.
A polygon of idealized excavation extent is used as a study area (a window).
<!-- Using a window closely approximating locations of the burials was considered,
e. g. buffer zone or convex/concave hull but simplified excavation polygon is
deemed the most useful because it preserves the borders and empty areas
of the original cemetery. -->

Overall structure of the cemeter is examined using
<!-- either density based approaches or -->
methods based on distances between neighboring burials. <!--[@osullivan2010, 121-154]-->
<!-- Density based approaches, for instance quadrat tests,
characterize the point pattern in terms of its *first-order properties* [@osullivan2010, 126].
The assumption underlying the *first-order effects* is that any point location is
independent on other point locations and different parameters influence its location.-->
These methods are based on distances between pairs and triples of neighbours and
characterize the point pattern based on *second- and third-order properties*
[@nakoinz2016, 130-135; @osullivan2010, 107].
These are in contrast to the assumption underlying *first-order effects* that
any point location is independent on other point locations and
defferent parameters influence its location [@osullivan2010, 126].
<!-- [@nakoinz2016, 135-144; @osullivan2010, 130-139] -->
In spatial analysis of settlement patterns individual locations of
settlements (points) can depend on a number of parameters, for instance
elevation, soil type etc.
In case of a grave placement in a burial ground we presume that the choice of
a graves location is either random or influenced predominantly by other (previous)
grave locations and
is largely independent on the outside parameters.
In other words *first-order properties* are of little to no importance in
our point of view.
The *second-order properties* and possibly *third-order properties*
are thus considered major factors in locating of the graves.
<!-- It is possible that the position of burials was dependent on parameters that were not
captured by the excavation, for example location of trees but that is impossible to test.-->

<!-- To characterize the pattern based on density, the quadrat test is employed.
It divides the area of interest into several cells of fixed sizes
(quadrats) and compares the count of events per quadrat with the count
expected under the null hypothesis that the underlying point process
is completely spatially random (approximated by a Poisson distribution). -->
We utilize *G*, *F*, Ripley's *K* [@ripley1976; @sayer2013] and *T* [@schladitz2000] functions.
These functions are based on nearest neighbour distances and are commonly used
to characterize point patterns in archaeology [@nakoinz2016, 135-144].
The *G* function is a metric of how close events are and summarises
cummulative frequency distribution of the nearest neighbour distances.
Because of its dependence on the intensity of events in the study area, it is
subject to changes in the size of the study area (window).
The *F* function, an empty-space function, is based on simulating a point pattern
and calculating of distances between random locations and the observed locations.
It is a measure of how far observed events are from any arbitrary location [@osullivan2010, 132-135].
Ripley's *K* function is in comparison with the *G* function resilient to
changes in the window size and works with all the proximal events up to a
certain threshold, not only the nearest neighbours [@nakoinz2016, 138].
We find it the most useful in most of archaeology use cases, cf. [@negre2018].
Last but not least, the *T* function is a third-order parallel to the *K* function [@schladitz2000].
For formulas and details on individual functions see the SI.

Given statistics for independent random process are estimated using Monte Carlo
simulation with 999 iterations.
Although smaller number of permutations should be sufficient to achieve
credible results
[@baddeley2016, 384] we follow the approach taken by @sayer2013 [78].
Ten most extreme realizations of the simulation are discarded as outliers so
the significance level *alpha* is 0.01 [@baddeley2016 410].

To explore frequency of neighboring burials falling into same categories we
explore several methods.
A modification of *K* function for multitype point patterns
does not take into account all of the
neighboring individuals in the point pattern but pairs of observations of the
same value within the given category are explored.
Same as *K* function is indicative of the distances at which clustering or
repulsive processes take place, the multitype *K* function shows these distances
for given pairs of levels within a given categorical variable.
The envelopes are constructed using the same procedure as described for other
functions.

To further establish the neighborhood structure we construct neighborhood graphs.
We use Delaunay triangulation and its subgraph, a Gabriel graph [@gabriel1969]
with burial locations as nodes (Fig. \@ref(fig:plan-graphs)).
The Delaunay triangulation is considered as a maximum number of neighbors per individual case
scenario while the Gabriel graph the opposite.
For each node, mean number of neighbors of a given category is determined and recorded.
To check how common the observed value is, a permutation test is performed
where the categories are randomly assigned while preserving the overall
proportion of given categories and the neighborhood structure determined by the
graph.
The observed values of neighbor categories are compared against distributions
derived from 999 permutations of the simulated data.

<!-- Percolation analysis -->
Another approach we consider for establishing neigborhood relations is by
constructing buffer zones around each of the burials.
To avoid determinism in designating the radius of the buffer zone around each of
the individuals we utilize percolation analysis.
Percolation analysis is a data centric method used to identify clusters of points based
on spatial separation measured by Euclidean distance [@maddison2020, 269].
The introduction of percolation analysis to archaeology is relatively recent,
nevertheless we consider its effectiveness and usefulness in studies with varied scales
well proven [@maddison2020].
Percolation analysis is based on city clustering algorithm [@rozenfeld2008] and
its generalization employing Euclidean distance [@rozenfeld2011].
Nodes being in a given distance threshold (radius) are assigned into a cluster.
This process is repeated for a set of thresholds while recording maximum
cluster sizes and possibly other statistics.
This allows to identify distances at which clusters exhibit rapid growth (absorbing
large number of nodes) and distances at which the clusters remain relatively stable.
For more details refer to @maddison2020 and associated R package [@schmidt2020].

By means of percolation analysis we identify threshold distances at which cluster
sizes are growing and clusters formed at these distances are further explored.
For each cluster number of individual bodies of a given category is recorded.
The same procedure is repeated for 999 resamples with the given
categories assigned at random with constant probabilities.
Observed numbers of occurences deviating by two standard deviations from the mean of the
resampled results are regarded as strong evidence of non-random patterns being
present in given clusters.
To achieve stability of results and eliminate observations occuring due to
chance in small clusters with low numbers of categories of interest,
the whole resampling procedure is repeated three times and only observations present in
all the iterations are reported with expected numbers of occurence and their
deviations from observed values summarized as mean values.

<!-- The size of the radius is based on the results of nearest neighbors analysis, where
at the given radius there is strong evidence of a clustering process (Fig. \@ref(fig:spat-funs)).
The randomness is again tested using a permutation test similar to the one
described for the neighborhood derived from the graphs.-->

<!-- software packages -->
Data is analyzed in an R environment for statistical computing and graphics
[@rcore] using functions of several different packages, for the complete list
see the supplementary information.

<!-- readr [@wickham2017a], dplyr [@wickham2016] and
reshape2 [@wickham2007] packages to manipulate the data.
Function daisy [@kaufman1990] from package cluster [@maechler2016] is used to
count distances between observations used in further analyses. -->

<!--
## Distance between burials

The input variables are of mixed data types, thus Gower's similarity
coefficient [@gower1971] is used in order to obtain a square distance matrix.
Gower's coefficient treats differently symmetric and asymmetric dichotomous
data, nominal and ordinal data and interval and ratio scaled data.
Advantage of this approach is that different kinds of variables can be analyzed
together and it is not necessary to transform the variables into same data
types, as it would be if other distance measure was used.
In our case, most of the variables concerning artifacts deposited within the
burials are asymmetric binary data, i. e. the presence of a trait means
similarity but absence does not imply similarity.
Continuous quantitative variables are treated as interval scaled.
Age categories are treated as ordinal variable and the rest of categorical
variables is treated as nominal data.
-->

<!-- ======================================================================= -->

<!-- results -->

# Results

## Artifact combinations

By counting co-occurences of analyzed variables on simulated data matrices and comparing
the distributions of the two studied statistics with the values calculated for
the original dataset we can establish whether the overall pattern of artefacts
distribution in the graves is non-random and further identify artefacts that deviate
from the random pattern.
The overall deviation from expected co-occurences, i.e. the *S* statistic
[@manly1996, 475] is significantly large at the 0.05 alpha level.
This suggests that there is strong evidence that the overall distribution of
artefacts in different graves is not random (see Fig. \@ref(fig:sstat)).

Examining distributions of individual artefacts it is found that at the 0.05
significance level only adzes, grinding tools, ceramic bowls and spondylus beads
are distributed randomly.
Close to the 0.05 alpha level are distributions of bone tools and ceramic bottles
(Fig. \@ref(fig:vstat)).
Among the grave goods that are found to be randomly distributed in the graves
across the burial ground are ochre pigment, lithics, pebbles, globular pots and
special types of pottery vessel shapes and spondylus pendants and *buckles*.
The relationships between artefact types frequently occuring and missing together
in graves are visualized in Fig. (\@ref(fig:coocnets)) with
clusters highlighted using fast-greedy algorithm [@newman2004] and strength of the
given association is expressed by the thickness of the edge connecting the nodes.
Artifacts most frequently present with other artifacts are with 5 links grinding tools and
ceramic bowls and with 4 links adzes, non-local lithic tools, ochre pigment and
spondylus beads.
The association is strongest between spondylus beads and (a) ceramic bottles and
(b) pigment; and adzes and (a) non-local lithics and (b) bone tools.
Artifacts that are frequently absent in the graves together are with 6 links
local lithic tools, globular pots and ceramic bowls and spondylus buckels with
5 and 4 links respectively.
The strongest antagonist relationships are between globular pots and (a) bowls and
(b) adzes.

## Spatial patterning

<!-- The quadrat test shows evidence for both random and clustered processes taking place with p values
under the 0.05 alpha level but the $\chi^2$ approximation is inaccurate due to small
expected counts (for the details see supplementary information).
Inconclusive result of the quadrat test is not suprising because as was noted
earlier, we do not expect *first-order* properties to play major roles in
placing of the individual burials. -->
Results of functions based on nearest neighbors are reported in Fig. \@ref(fig:spat-funs).
Both *G* and  *F* functions do slightly deviate from the curve expected
under independent random process
but neither traverses the boundaries of the envelope.
Thus the cummulative nearest-neighbor distance (*G*) as well as the
empty-space function *F* are well approximated by
random process and do not show evidence for spatial clustering nor repelent functions
taking place.
On the contrary
the *K* and *T* functions show clear deviations from the expected curve and both exceed
the boundaries of given envelopes at distances around or smaller than 5 meters.
<!-- in case of *K* and slightly more in case of *T* function.-->
This means that numbers of nearest neighbors at this and larger distances is
significantly higher than we would expect under complete spatial randomness.
The overall point pattern thus shows clear signs of clustering process taking place.

This means that the locations of new burials were not chosen at random
but were affected by the positions of previous burials.
The pattern begins to cluster at the distance of around 5 meters, meaning
that beginning at this distance, the probability of finding a neighboring burial is
much higher than it would be if the burials were placed randomly in space.
The community burying their dead at the cemetery of Vedrovice *Široká u lesa*
did so in an organized manner taking into account previous burials at the same place,
meaning their positions were known, probably marked on the surface somehow.
This would imply existence of non-perishable marks on the surface of the cemetery and/or their
maintenance from time to time.

Results of percolation analysis reveal more detailed view of the clustering process
taking place, see Fig. \@ref(fig:perc-distance).
The distances where relative stability of clusters emerges are labeled.
The curve reveals a steady growth up until
radius of 5 meters and then a large skip towards a single cluster
between 5.5 and 6.5 meters.
This point of cluster growth is revealed by the point pattern analysis as well.
The clusters at some of the distances are shown in Fig. \@ref(fig:perc-plan).

## Neighborhood structure

The results of multitype *K* function show deviations from random processes for
most of the observed categories in Fig. \@ref(fig:pp-crossk).
The distances at which these deviations start occuring are in most cases rather large,
well over 10 meters and considering the fact that at the 10.4 meters radius
around individual points the percolation analysis proved existence
of single large cluster encompassing almost all of the points, these
observations correspond with the deviations observed for single *K* function.
The only notable deviation we consider worth mentioning is in case of age
category in pairs of juveniles and in sex category between pairs of non-adults.
Here the deviation from random processes is observed at smaller distances than
in case of other variables at distances around 5 meters.
This proves tighter bound cluster of juveniles in south-eastern part of the cemetery.

In case of neighborhood structure derived from Delaunay triangulation (maximum
number of neighbors scenario) and
Gabriel graph (minimum number of neighbors scenario, Fig. \@ref(fig:plan-graphs))
most of the observed relationships are not significantly different from those
derived by randomization experiment.
In the Delaunay triangulation the significant relations are observed in two cases
between (1) pairs of non-adults and (2) female-sexed bodies and non-adults.
The first observation can be again given into relation with higher concentration of
non-adults in south-eastern space of the burial ground.
In regard to sex category Gabriel graph shows an association
between female-sexed and male-sexed bodies.
In only one case are the observations from both of the graphs significant and
that is between pairs of bodies of mature age category.

In case of clusters at different percolation distances
there are some deviations from random number of
occurences of given categories at different scales.
Regarding the determination of sex of the bodies, at the threshold of 3.9 meters
we observe that in the largest cluster (nr. 3, Fig. \@ref(fig:perc-plan))
there are more non adults and less male-sexed bodies and in the second largest
cluster (nr. 10) there are slightly more male-sexed bodies than would
be expected under random process.
The overrepresentation of non-adult bodies is repeated in the following
threshold distance of 4.6 meters in the largest cluster (nr. 3).
Last but not least, in the cluster nr. 8 (north of the cemetery) at the distance
of 5.5 meters female-sexed bodies are more frequent than we would expect.
The age categories correspond to the previous obsevations where
most evident is the overrepresentaion of juvenile bodies at the 3.9 meters threshold in
cluster nr. 3.
Exploring the origin of the bodies based on Strontium isotope signature we
find that cluster in the south-western part at threshold distances 3.9 and 5.5 meters
(nr. 10 and 5 respectively) consists of less non-local bodies than would be expected.
Lastly at the distance radius 2.8 meters there are more local bodies in the cenral
cluster of the cemetery (nr. 8).

<!-- Methods of neighborhood relationship construction examined include
distance based approaches where all the neighbors that fall into a given buffer are considered,
the other two are based on neighborhood graphs, the Delaunay triangulation as a
maximum number of neighbors scenario and a Gabriel graph as a minimal scenario. -->
<!--
For relationships (see Fig. \@ref(fig:nb-sex) and Tab. \@ref(tab:nb-tab-sex))
of male-sexed bodies, none of the methods showed any significant
differences between expected and observed number of neighbors of a given sex.
The relationship with female-sexed bodies is rather extreme for all of the methods employed,
but never significant ($p > 0.07$).
In case of bodies unsexed due to early age (non-adults), the neigborhood based
on buffer zones show a signficant realtionship with other non-adults ($p \approx 0.04$).
Finally burials with female-sexed bodies show a significantly higher number of
burials with male-sexed bodies in both the methods based on neighborhood networks
(Delaunay triangulation $p \approx 0.02$ and Gabriel graph $p \approx 0.03$).
Two further relationships showed non-random deviations.
In case of Delaunay trinagulation for burials with bodies with indeterminate sex with
non-adults ($p \approx 0.02$) and in case of Gabriel graph between non-adults and
indeterminate category $p \approx 0.02$.
-->

<!-- ```{r t-vstat} -->
<!-- vstat <- read_csv(here("data/temp", "v_statistics.csv")) %>% -->
<!--     select(full, v, p) %>% -->
<!--     mutate(v = round(v, 2), p = round(p, 2)) %>% -->
<!--     rename("variable" = full, "v statistic" = v, "p value" = p) -->

<!-- knitr::kable( -->
<!--   vstat, -->
<!--   caption = 'Fuuuuu' -->
<!-- ) -->
<!-- ``` -->

<!-- ======================================================================= -->

# Discussion

<!-- ======================================================================= -->
<!-- FIGURES --><!-- FIGURES --><!-- FIGURES --><!-- FIGURES --><!-- FIGURES -->

```{r sstat, fig.cap = "Distribution of simulated and observed *S* statistic"}
knitr::include_graphics(here("plots/s_statistic.pdf"))
```

```{r vstat, fig.cap = "Simulated and observed *V* statistics for different variables"}
knitr::include_graphics(here("plots", "v_values.pdf"))
```

```{r coocnets, fig.cap = "Networks of variables frequently present together and exluding each other"}
knitr::include_graphics(here("plots", "cooc_networks.pdf"))
```

<!-- ```{r plan-graphs, fig.cap = "Neighborhood graphs: Delaunay triangulation and Gabriel graph"} -->
<!-- knitr::include_graphics(here("plots/plan_gabriel.pdf")) -->
<!-- ``` -->

<!-- ```{r label-spat-funs} -->
<!-- label_spat_funs <- paste("Functions *G*, *F*, *K* and *T*.", -->
<!-- "*Dashed line* -- theoretical curve under complete spatial randomness, *solid line* -- empirical curve, *gray area* -- envelope.") -->
<!-- ``` -->
<!-- ```{r spat-funs, fig.cap=label_spat_funs} -->
<!-- knitr::include_graphics(here("plots/pointprocess_fun.pdf")) -->
<!-- ``` -->

<!-- ```{r label-nb-sex} -->
<!-- label_nb_sex <- paste("Mean neighbors in a given sex category.", -->
<!-- "*Gray area* -- density estimation for the simulated data, *vertical line* -- observed value.", -->
<!-- "*F* -- female-sexed body, *M* -- male-sexed body, *n. a.* -- non adult, *ind.* -- indetermined sex.") -->
<!-- ``` -->
<!-- ```{r nb-sex, fig.cap=label_nb_sex} -->
<!-- knitr::include_graphics(here("plots/nb_sex.pdf")) -->
<!-- ``` -->

<!-- ======================================================================= -->
<!-- TABLES --><!-- TABLES --><!-- TABLES --><!-- TABLES --><!-- TABLES -->
<!-- ```{r nb-tab-sex} -->
<!-- tab <- read_csv(here("data/temp/nb_sex_p.csv")) %>% -->
<!--          mutate(relation = str_c(from, to, sep = " - ")) %>% -->
<!--   select(relation, everything(), -from, -to) -->

<!-- tnames <- c("relation", rep(c("mean", "p"), 3)) -->

<!-- knitr::kable(tab, booktabs = TRUE, col.names = tnames, linesep = "", caption = "Results...") %>% -->
<!--   kableExtra::add_header_above(c(" ", "Buffer" = 2, "Delaunay" = 2, "Gabriel" = 2)) -->
<!-- ``` -->

<!-- ======================================================================= -->

# References