### 2021-06-03 text/article_vedrovice.Rmd
update text after proofreading

---
title           : "Spatial Patterns and Grave Goods Differences at the Cemetery of Vedrovice (Czech Republic): A Resampling Approach to Identity Markers in the Early Neolithic"
author          : "Petr Pajdla"
date            : "`r format(Sys.time(), '%d %B, %Y')`"
abstract        :
  "The spatial pattern and grave good associations at the cemetery of Vedrovice - Široká u Lesa are explored using computationally intensive formal methods. Resampling approaches to both intra-site spatial arrangement and the distribution of grave goods are explored in relation to layers of identity represented by different buried body attributes – sex, age, geographical origin and their combinations. We found that polished stone adzes, grinding tools, ceramic bowls and spondylus beads occur among non-randomly distributed grave goods. As a result of the investigation of differences among the non-randomly distributed artifacts and attributes of burials, we conclude that the most significant are dichotomies between genders, accentuated by both origin and age. Spatial analysis shows clear trends towards clustering between burials, and the emerging clusters are further explored for similarities in burial attributes."
output:
    bookdown::word_document2:
        latex: pdflatex
        toc: false
        fig_caption: true
        number_sections: false
        keep_tex: false
bibliography    : "bib_vedrovice.bib"
csl             : "apa.csl"
header-includes:
  - \usepackage[left]{lineno}
  - \linenumbers
---

<!-- note: length - approx. 10 - 12 pages (no bibliography, tables, figures) -->
<!-- i.e. 18 000 - 22 000 characters -->

```{r init, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, message = FALSE, warning = FALSE)
# library(tidyverse)
library(here)
```

<!-- ======================================================================= -->

# Introduction

The Cemetery of Vedrovice is one of the earliest and most intensively studied
*Linearbandkeramik* (LBK) cemeteries.
Recent articles use evidence from the cemetery in a broader discussion
of post-marital residence patterns [@hrncir2020],
sexual inequalities and gender in the Neolithic [@masclanslatorre2020; @bickle2020].
Here we focus on an exploration of the relationship between three proposed *layers* of
identity, biological sex and age of the body determined from skeletal remains
by anthropological methods and the origin of the body based on the strontium isotope
signal.
We explore whether the structure observed in these three categories is
mirrored in the grave good associations and spatial organization within the
cemetery.
The analysis is set into a broader discussion of gender relationships and identity
in Neolithic communities.

<!-- view of gender in the Neolithic, archaeological context -->
@robb2017 [p. 128] noted a lack of debate in archaeology on gender for the
European Neolithic compared to other periods.
The authors present gender in the Neolithic as being
of a very different nature than gender in later prehistory.
It is argued that gender is not clearly expressed and
biological differences between sexes are not transformed into binary cultural
categories [@robb2017, p. 140--142].
Considering the Neolithic burial evidence across Europe, authors find that it
"shows very little gender distinction; when it does, it is rarely a clear-cut
binary categorization but a matter of degree and overlap, sometimes polythetically
combined with factors such as age and localness" [@robb2017, p. 138].
In response, @bickle2020 [p. 14] argues that "biological sex may not have defined gender
during the LBK" although the "sexed bodies did matter".

<!-- We focus on relations between burials and bodies in terms of biological
sex, age and origin, exceptionality of grave goods and location within the
cemetery.-->
We focus mainly on investigating non-random patterns in artifact
co-occurrences and spatial arrangement in the cemetery in relation to the
proposed *layers* of Neolithic identity, i.e., sex, age and localness of the body.
Our question is thus how grave goods relate to the aforementioned
categories and whether there is a consistent relationship between
these categories forming the *identity* and spatial
arrangement of the cemetery.

<!-- gender relations in LBK -->
<!-- Although gender relations in the LBK might not be -->

<!-- resampling and computer-intensive techniques-->
The chosen approach is a quantitative one embracing resampling techniques and
computer-intensive methods in general.
The resampling techniques were introduced in mortuary studies by
@manly1996 but the application is still limited [@sosna2008, p. 343].
As noted by @sosna2008, the most important benefit of resampling is its
ability to distinguish random and non-random patterns.
This is crucial for mortuary analysis because the studied samples, including
the one presented here, are usually small in size and identifying which
observed patterns are caused by mere chance associated with small numbers
and which are actually present is problematic.

<!-- ### Questions
* Is there evidence that certain artifacts covary,
either positively or negatively?
Do certain artifacts tend to occur together, or are certain combinations
absent or rare?
* How can the presence and absence data best display which burials
contain similar artifacts,
and which artifacts have similar distributions over burials?
* *How can the data be analyzed so as to show most clearly the relationships*
*between the different* **burials** in terms of their **contents**?
* *How can the data be analyzed so as to show most clearly the relationship*
*between the different* **artifacts** in terms of their
**distributions in graves**?
* How are grave good assemblages related to age and sex?
(... variations in burial rite coinciding with age or sex of the deceased ...) -->

<!-- ======================================================================= -->

# Material and methods

<!-- site -->
The Vedrovice site is situated in south Moravia (Czech Republic).
The earliest known finds and archaeological explorations here are from
the 19th century, but excavations were conducted in the second half of 20th century.
There are two known burial areas, *Za Dvorem* and the larger *Široká u Lesa* and
a settlement area.
For more details on the site, see @masclanslatorre2020 or the
original publication by @podborsky2002.
The material analyzed here comes from the *Široká u lesa* cemetery in Vedrovice.

<!-- sample (burials, artifacts) -->
Originally, this burial ground consisted of at least 108 burials.
Some of the burials were disturbed by later settlement and burial pits in the
area of the cemetery, amateur excavations during the 19th century and later
agricultural activities.
These burials are excluded from most of the analysis due to missing and/or
incomplete data and only a sample of 78 well-preserved LBK burials is analyzed here.
Most of the data was obtained from the original publication [@podborsky2002],
and the anthropological determination of the sex and age of the bodies was corrected based on an
analysis by @dockalova2008. Carbon, nitrogen and strontium isotope
ratios were collected in @whittle2013 and @richards2008.
Further details on the use and function of the artifacts were gathered from
@masclanslatorre2020.
Information on the localness of various bodies was adapted from @hrncir2020.

Each burial was originally characterized by 20 variables recording the number of
artifacts present in the graves.
Some of  these original variables are omitted in the analysis due to their rare occurrence;
for instance, a spondylus arm ring is present in only a single burial,
rendering it useless in establishing relationships between pairs of burials.
Only grave goods occurring in more than 5% of the graves are used in the analysis;
this step reduces the number of variables to 14.
These include variables noting quantities of various types of pottery vessels in
the graves, counts of chipped stone tools made of either local or non-local raw materials,
the presence of polished and beveled stone adzes, grinding slabs, pebbles,
bone tools and different kinds of personal ornaments.
The personal ornaments include beads made of spondylus shells,
spondylus pendants and oval shaped buckles.
Last but not least, the presence of ocher pigment in the grave is also noted.
Quantities of various *grave goods* are simplified into a binary matrix indicating
the presence or absence of a given trait.

Another category of variables is derived from retrieved human skeletal remains.
These include biological sex, age simplified into age groups and an indicator
variable as to whether the strontium signal in the remains is of local origin.
The variable representing the recorded sex is simplified into the following
categories: female-sexed body (abbreviated *F*), male-sexed body (*M*),
non-adult (*n. a.*) and undetermined sex (*ind.*).
The age of the body is also simplified into three categories: juveniles (*juv.*)
comprised of the groups *Infans* and *Juvenis*, adults (*ad.*, all bodies aged as *Adultus*)
and mature bodies (*mat.*, anthropological categories *Adultus-Maturus*, *Maturus* and
*Senilis*) and bodies of an undetermined age (*ind.*).
The last category is the indicator for the origin of the buried person.
The local strontium range is defined as two standard deviations from the mean of Sr ranges of
bodies with an age category determined as *Infans*.
Bodies falling within this range are labeled as local, outside the range as non-local;
where data on Sr values is missing, origin is not determined (*ind.*).

As mentioned above, we consider these three variables -- sex, age and origin -- to be indicative of
an individual's identity in the Neolithic.
To explore the relationships between the variables, we use them both as individual
categories and combined into the following indicator variables -- (1) sex/age;
(2) origin/sex; (3) origin/age; and (4) origin/sex/age.
Due to frequent missing or undetermined values, we decided to limit the studied
sample only to bodies with all of the categories determined (and only categories
that occur in more than 5% of burials).

<!-- methods -->

## Randomness of artifact combinations

<!--
This study takes advantage of computer-intensive resam-
pling techniques that are based on simulations of the behavior
of statistical models through the generation of random samples
(Baxter, 2003; Efron and Tibshirani, 1993; Manly, 1991). The
strength of these techniques lies in the potential to effectively
distinguish between random and non-random patterns and to
test differences between samples without making assumptions
about the population from which these samples were drawn
(Drennan and Peterson, 2004). These techniques proved to
be very effective for the analysis of patterns in cemeteries
(Manly, 1996).
@sosna2008 -->

To determine the non-randomness of various artifact combinations in burials, we conducted a
permutation test on the presence/absence data matrix.
The method based on generalized Monte Carlo tests was
introduced into archaeology by @manly1996 and adapted by @sosna2008.
Our approach is based on both of these works.
The original co-occurrence matrix was permuted 9,999 times with constant
row and column totals, i.e., both the count of artifacts in individual graves
and the sum of the given type of artifact across the burials are fixed.

To examine which associations of variables, i.e., artifacts, are infrequent or even not present
and which tend to appear together more often than it would be expected by random chance,
the co-occurrences of observed and simulated data
are summarized using two statistics suggested by @manly1996 [475].
The *S* statistic is a measure of overall deviation between the expected and observed
number of artifact occurrences, while
the *v* statistic is a measure of difference between co-occurrences of
individual artifacts with other artifacts.

To assess whether differences between artifacts that are non-randomly distributed
are reflected in relationships between groups based on sex, age, origin and their combinations,
we calculate mean Euclidean distance (ED) between pairs of burials in multidimensional
space settings defined by the non-randomly distributed artifacts.
The observed mean ED is compared to the mean ED distribution  derived by random
reshuffling of the category labels between the burials 9,999 times with the constant
probabilities of appearance of the given category label.
<!-- To infer whether the observed and randomly created mean ED
distributions are samples of the same population we use non-parametric
Kruskal-Wallis test [@kruskal1952].-->
For each of the categories, the deviation of the mean ED from its randomly generated
distributions is expressed as the fraction of random values of ED larger than
the one observed.
The deviation is deemed significant if less than 5 % of
random values are larger than the observed value.
The process is a variation of the one described for resampling in the case of
artifact combinations, and a similar approach is taken by @sosna2008 [346-347].

<!--
To assess relations between groups based on sex, age and origin
statistics introduced by @sosna2008 are used.-->

<!-- ## Exceptionality index -->

<!-- Exceptionality index (*EI*) is used in order to induce relative exceptionality -->
<!-- of individual burials based on their attributes. -->
<!-- This approach is independent on a prior selection of attributes of importance -->
<!-- or *prestige*. -->
<!-- It was defined by @demjan2015[367] based on a Root Mean Square (*quadratic mean*) of -->
<!-- PCA factor scores: -->
<!-- $$ EI_{attr_g}=\sqrt{\frac{1}{n}\sum_{i=1}^{n}s_i^2} $$ -->
<!-- where $s_i$ are factor scores of individual graves. The mean value is -->
<!-- normalized to range $[0-1]$: -->

<!-- $$ EI_{attr_g}=\frac{EI_{attr_g}}{max~EI_{attr_g}} $$ -->

<!-- spatial analysis -->
## Spatial analysis

Spatial patterning in prehistoric burial grounds is often examined to find
any internal structure, arrangement or possible grouping.
We approach the task as a point pattern analysis problem given that
a burial (event) is simplified as its centroid location in two dimensional space.
A polygon of the idealized excavation extent is used as the study area (a window).
<!-- Using a window closely approximating locations of the burials was considered,
e. g. buffer zone or convex/concave hull but simplified excavation polygon is
deemed the most useful because it preserves the borders and empty areas
of the original cemetery. -->

<!-- Density based approaches, for instance quadrat tests,
characterize the point pattern in terms of its *first-order properties* [@osullivan2010, 126].
The assumption underlying the *first-order effects* is that any point location is
independent on other point locations and different parameters influence its location.-->
The overall structure of the cemetery is examined using
methods utilizing distances between neighboring burials.<!--[@osullivan2010, 121-154]-->
These methods are based on distances between pairs and threesomes of neighbors and
characterize the point pattern based on *second- and third-order properties*
[@nakoinz2016, 130-135; @osullivan2010, 107].
These are in contrast to the assumption underlying *first-order effects* that
any point location is independent of other point locations and
different parameters influence its location [@osullivan2010, 126].
In the spatial analysis of settlement patterns, individual locations of
settlements (points) can depend on a number of parameters, e.g.,
elevation, soil type, etc.
In the case of a grave placement in a burial ground, we presume that the choice of
the grave location is either random or influenced predominantly by other (previous)
grave locations and
is largely independent of outside parameters.
In other words, *first-order properties* are of little to no importance in
our point of view.
The *second-order properties* and possibly *third-order properties*
are thus considered major factors in the location of the graves.
<!-- It is possible that the position of burials was dependent on parameters that were not
captured by the excavation, for example location of trees but that is impossible to test.-->

<!-- To characterize the pattern based on density, the quadrat test is employed.
It divides the area of interest into several cells of fixed sizes
(quadrats) and compares the count of events per quadrat with the count
expected under the null hypothesis that the underlying point process
is completely spatially random (approximated by a Poisson distribution). -->
We utilize *G*, *F*, Ripley's *K* [@ripley1976; @sayer2013] and *T* [@schladitz2000] functions.
These functions are based on the nearest neighbor distances and are commonly used
to characterize point patterns in archaeology [@nakoinz2016, 135-144].
The *G* function is a metric of how close events are and summarizes
cumulative frequency distribution of the nearest neighbor distances.
Because of its dependence on the intensity of events in the study area, it is
subject to changes in the size of the study area (window).
The *F* function, an empty-space function, is based on simulating a point pattern
and calculating the distances between random locations and the observed locations.
It is a measure of how far observed events are from any arbitrary location [@osullivan2010, 132-135].
Compared to the *G* function, Ripley's *K* function is resilient to
changes in the window size and works with all the proximal events up to a
certain threshold, not only the nearest neighbors [@nakoinz2016, 138].
We find it the most useful in most archaeology cases, cf. @negre2018.
Last but not least, the *T* function is a third-order parallel to the *K* function [@schladitz2000].

The given statistics for the independent random process are estimated using Monte Carlo
simulation with 999 iterations.
Although a smaller number of permutations should be sufficient to achieve
credible results
[@baddeley2016, 384], we follow the approach taken by @sayer2013 [78].
The ten most extreme realizations of the simulation are discarded as outliers, so
the significance level is 0.01 [cf. @baddeley2016 410].
The remaining realizations of given point processes are used to construct envelopes
around function curves derived from observed data.

To explore the frequency of neighboring burials falling into the same categories, we
explore several methods.
A modification of the *K* function for a multitype point pattern
does not take into account all of the
neighboring individuals in the point pattern but pairs of observations of the
same value within the given category are explored.
In the same way that the *K* function is indicative of the distances at which clustering or
repulsive processes take place, the multitype *K* function shows these distances
for the given pairs of levels within a given categorical variable.
The envelopes are constructed using the same procedure as described for other
functions.

To further explore the neighborhood structure, we construct neighborhood graphs.
We use Delaunay triangulation and its subgraph, a Gabriel graph [@gabriel1969]
with burial locations as nodes (Fig. \@ref(fig:plan-graphs)).
The Delaunay triangulation is regarded as the maximum number of neighbors per individual case
scenario, while the Gabriel graph is the opposite.
For each node, the mean number of neighbors of a given category is determined and recorded.
To check how common the observed value is, a permutation test is performed
where the categories are randomly assigned while preserving the overall
proportion of given categories and the neighborhood structure determined by the
graph.
The observed values of neighbor categories are compared to distributions
derived from 999 permutations of the simulated data.

<!-- Percolation analysis -->
Another approach we consider for establishing neighborhood relationships is
constructing buffer zones around each of the burials.
To avoid determinism in designating the radius of the buffer zone around each of
the individuals, we utilize percolation analysis,
a data centric method used to identify clusters of points based
on spatial separation measured by Euclidean distance [@maddison2020, 269].
Although the introduction of percolation analysis in archaeology is relatively recent,
we consider its effectiveness and usefulness in studies with varied scales
well-proven [@maddison2020].
Percolation analysis is based on a city clustering algorithm [@rozenfeld2008] and
its generalization employing Euclidean distance [@rozenfeld2011].
Points falling within a given distance threshold (radius) are added to a cluster.
This process is repeated for a set of thresholds while recording maximum
cluster sizes and possibly other statistics.
This makes it possible to identify distances at which clusters exhibit rapid growth (absorbing
a large number of nodes) and distances at which the clusters remain relatively stable.
For more details, refer to @maddison2020 and the associated R package [@schmidt2020].

By means of percolation analysis, we identify threshold distances at which maximum cluster
size is growing, and clusters formed at these distances are further explored.
For each cluster, the number of individual bodies of a given category is recorded.
The same procedure is repeated for 999 resamples with the given
categories assigned at random with constant probabilities.
The observed numbers of occurrences deviating by more than two standard deviations from the mean of the
resampled results are regarded as strong evidence of non-random patterns being
present in given clusters.
To achieve the stability of results and eliminate observations occurring by
chance in small clusters with low numbers of categories of interest,
the whole resampling procedure is repeated three times and only observations present in
all the iterations are reported with the expected numbers of occurrence and their
deviations from observed values summarized as mean values.

<!-- The size of the radius is based on the results of nearest neighbors analysis, where
at the given radius there is strong evidence of a clustering process (Fig. \@ref(fig:pp-funs)).
The randomness is again tested using a permutation test similar to the one
described for the neighborhood derived from the graphs.-->

<!-- software packages -->
Data is analyzed in an R environment for statistical computing and graphics
[@rcoreteam2021] using functions of several different packages; for the complete list,
see the supplementary information.

<!-- readr [@wickham2017a], dplyr [@wickham2016] and
reshape2 [@wickham2007] packages to manipulate the data.
Function daisy [@kaufman1990] from package cluster [@maechler2016] is used to
count distances between observations used in further analyses. -->

<!--
## Distance between burials

The input variables are of mixed data types, thus Gower's similarity
coefficient [@gower1971] is used in order to obtain a square distance matrix.
Gower's coefficient treats differently symmetric and asymmetric dichotomous
data, nominal and ordinal data and interval and ratio scaled data.
Advantage of this approach is that different kinds of variables can be analyzed
together and it is not necessary to transform the variables into same data
types, as it would be if other distance measure was used.
In our case, most of the variables concerning artifacts deposited within the
burials are asymmetric binary data, i. e. the presence of a trait means
similarity but absence does not imply similarity.
Continuous quantitative variables are treated as interval scaled.
Age categories are treated as ordinal variable and the rest of categorical
variables is treated as nominal data.
-->

<!-- ======================================================================= -->

<!-- results -->

# Results

## Artifact combinations

By counting the co-occurrences of analyzed variables on simulated data matrices and comparing
the distributions of the statistics proposed by @manly1996 with the values calculated for
the original data set, we can establish whether the overall pattern of artifact
distribution in the graves is non-random and further identify artifacts that deviate
from the random pattern.
The overall deviation from expected co-occurrences, i.e., the *S* statistic
[@manly1996, 475] is significantly large at the 0.05 alpha level.
This suggests that there is strong evidence that the overall distribution of
artifacts in different graves is not random. <!--(see Fig. \@ref(fig:sstat))-->

Examining distributions of individual artifacts, we find that at the 0.05
significance level only adzes, grinding tools, ceramic bowls and spondylus beads
are distributed non-randomly.
Close to the 0.05 alpha level are distributions of bone tools and ceramic bottles
(Fig. \@ref(fig:cooc-vstat)).
Among the grave goods found to be randomly distributed in the graves
across the burial ground are ocher pigment, lithics, pebbles, globular pots and
special types of pottery vessel shapes, spondylus pendants and buckles.

The relationships between artifact types frequently occurring and missing together
in graves are visualized in Fig. \@ref(fig:cooc-nets) with
clusters established using a fast-greedy algorithm [@newman2004] and the strength of the
given association expressed by the thickness of the edge connecting the nodes.
Artifacts most frequently present with other artifacts are grinding tools and
ceramic bowls with five links and adzes, non-local lithic tools, ocher pigment and
spondylus beads with four links.
The association is strongest between spondylus beads and (a) ceramic bottles and
(b) pigment; and adzes and (a) non-local lithics and (b) bone tools.
Artifacts frequently absent with other artifacts in the graves are  
globular pots, local lithic tools, ceramic bowls and spondylus buckles
with seven to four links respectively.
The strongest antagonist relationships are between globular pots and (a) bowls and
(b) adzes.

Artifacts non-randomly distributed across the cemetery are further used to
check whether significant differences exist between their occurrences in
different groups defined based on origin, sex, age and their combinations.
Exploring single attributes, the only differences are in the categorization to
differently sexed bodies; there are significant differences
between male- and female-sexed bodies.
In the case of age or origin, no significant differences are observed.
If we combine sex and age categories, the distinction between juvenile, adult and mature
female-sexed bodies and adult male-sexed bodies is evident.
Another observed dichotomy is present between adult males and juveniles (non-adults).
The same relationships are essentially present in the case of combinations of sex and origin attributes.
Local male-sexed bodies have significantly different combinations of non-randomly
distributed artifacts from both local and non-local female-sexed bodies and non-adults.
If all the categories are combined the divisions are even more evident.
Local male-sexed bodies in the *Adultus* age category deviate in the artifact distributions
from adult local and non-local females, *Maturus/Senilis* non-local female-sexed bodies
and juveniles.

## Spatial patterning

<!-- The quadrat test shows evidence for both random and clustered processes taking place with p values
under the 0.05 alpha level but the $\chi^2$ approximation is inaccurate due to small
expected counts (for the details see supplementary information).
Inconclusive result of the quadrat test is not surprising because as was noted
earlier, we do not expect *first-order* properties to play major roles in
placing of the individual burials. -->
The results of functions based on nearest neighbors are reported in Fig. \@ref(fig:pp-funs).
Both *G* and  *F* functions slightly deviate from the curve expected
under an independent random process,
but neither crosses the boundaries of the envelope.
Thus, the cumulative nearest-neighbor distance (*G*) as well as the
empty-space function (*F*) are well approximated by
the random process and do not show evidence for spatial clustering or repellent functions
taking place.
On the contrary,
the *K* and *T* functions show clear deviations from the expected curve and both exceed
the boundaries of the given envelopes at distances around or smaller than 5 meters.
<!-- in case of *K* and slightly more in case of *T* function.-->
This means that the number of nearest neighbors at this and larger distances is
significantly higher than we would expect under complete spatial randomness.
The overall point pattern thus shows a clear signs of clustering process taking place.

This suggests that the locations for new burials were not chosen at random
but were influenced by the positions of previous burials.
The pattern begins to cluster at a distance of around 5 meters, i.e.,
beginning at this distance, the probability of finding a neighboring burial is
much higher than it would be if the burials were placed randomly in space.
The community burying their dead at the *Široká u lesa* cemetery in Vedrovice
did so in an organized manner taking into account previous burials at the same place,
meaning their positions were known, probably marked somehow on the surface.
This would imply the existence of non-perishable marks on the surface of the cemetery and/or their maintenance from time to time.

The results of a percolation analysis reveal a more detailed view of the clustering process
taking place; see Fig. \@ref(fig:perc-distance).
The distances at which the relative stability of clusters emerges are labeled.
The curve reveals a steady growth up until a
radius of 5 meters and then a large skip towards a single cluster
between 5.5 and 6.5 meters.
This point of cluster growth is revealed by the point pattern analysis as well.
The clusters at chosen distances are shown in Fig. \@ref(fig:perc-plan).

## Neighborhood structure

The results of the multitype *K* function show deviations from random processes for
most of the observed categories in Fig. \@ref(fig:pp-crossk).
The distances at which these deviations start occurring are in most cases rather large,
well over 10 meters and, considering the fact that at the 10.4 meters radius
around individual points the percolation analysis proved the existence
of a single large cluster encompassing almost all of the points, these
observations correspond to the deviations observed for the single *K* function.
The only notable deviation we consider noteworthy is in the case of age
category in pairs of juveniles and in the sex category between pairs of non-adults.
Here the deviation from random processes is observed at smaller distances than
in the case of other variables at distances around 5 meters.
This proves a tighter bound cluster of juveniles in the southeast part of the cemetery.

Multitype *K* for the combination of age and sex categories suggests a clustering process
between bodies determined as adult females and juveniles and vice versa.
The relationship between pairs of non-adults/juveniles is also confirmed.
Several other observed deviations occur at large distances and the explanation for
these is the same as in the case of multitype *K* for single categories.

In the case of neighborhood structure derived from Delaunay triangulation (maximum
number of neighbors scenario) and
Gabriel graph (minimum number of neighbors scenario, Fig. \@ref(fig:plan-graphs)),
most of the observed relationships are not significantly different from those
derived by randomization experiment.
In the Delaunay triangulation the significant relationships are observed in two cases
between (1) pairs of non-adults and (2) female-sexed bodies and non-adults.
The first observation can again be connected with a higher concentration of
non-adults in the southeast space of the burial ground.
In regard to the sex category, the Gabriel graph shows an association
between female-sexed and male-sexed bodies.
In only one case are the observations from both of the graphs significant, and
that is among pairs of bodies in the mature age category.

The combinations of categories in the Delaunay graph show significant relationships for
pairs of bodies labeled as *Maturus/Senilis* category females with
(1) juvenile females and with (2) unsexed juveniles.
The relationship between pairs of juvenile/non-adult bodies is again confirmed.
The Gabriel graph in the sex/age category shows relationships between (1) *Adultus* female and male bodies;
(2) *Maturus* female and male bodies; and (3) pairs of *Maturus* female bodies.
The origin/sex combination of categories produces a significant deviation from
expected frequencies for (1) non-local female bodies and local non-adults in
Delaunay triangulation and in a Gabriel graph for (1) non-local female and local male
bodies and vice versa.
The origin/age combination shows significant departures only in the Gabriel graph for
local *Adultus* bodies neighboring with (1) non-local *Maturus* bodies and (2) local juvenile bodies.
Last but not least, incorporating all of the categories into the networks, we observe
significant results in both of the graphs for non-local *Maturus* female bodies and
local juvenile (non-adult) bodies and in the case of the Gabriel graph a relationship
between non-local *Adultus* female and local *Maturus* male bodies.

Clusters formed at different percolation distances
show the following deviations from the expected number of
occurrences of given categories at different scales.
Regarding the determination of the sex of the bodies, at the threshold of 3.9 meters
we observe that in the largest cluster (no. 3, Fig. \@ref(fig:perc-plan))
there are more non-adults and less male-sexed bodies and in the second largest
cluster (no. 10) there are slightly more male-sexed bodies than would
be expected under a random process.
The over-representation of non-adult bodies is repeated in the following
threshold distance of 4.6 meters in the largest cluster (no. 3).
Last but not least, in cluster no. 8 (north of the cemetery) at the distance
of 5.5 meters, female-sexed bodies are more frequent than we would expect.
The age categories correspond to the previous observations where
the over-representation of juvenile bodies at the 3.9-meter threshold in
cluster no. 3 is most evident.
Exploring the origin of the bodies based on the strontium isotope signature, we
find that the cluster in the southwest part at threshold distances of 3.9 and 5.5 meters
(no. 10 and 5 respectively) consists of less non-local bodies than would be expected.
Finally, at a radius of 2.8 meters there are more local bodies in the central
cluster of the cemetery (no. 8).

In the combined categories most of the already observed patterns are repeated and
highlighted.
In the largest cluster across percolation thresholds at 3.9 and 4.6 meters
(no. 3), there is an abundance of (1) local non-adults/juveniles and (2) non-local
female-sexed bodies of the *Maturus* age category and, on the other hand,
lower numbers of (3) juvenile female-sexed bodies.
The central cluster across the distances of 2.8, 3.9 and 4.6 meters (no. 8, 10 and 7 respectively)
shows higher numbers of local male- and female-sexed bodies of the *Adultus* category
and, in the case of female bodies, the *Maturus* age categories.
The southwest cluster (no. 10, 7 and 5) at percolation distances 3.9, 4.6 and 5.5 meters
is clearly dominated by local male-sexed bodies of the *Maturus* age category.
Last but not least, in the case of all the categories combined together, a cluster in
the western part of the cemetery at the percolation threshold of 4.6 and 5.5 meters (no. 8 and 6 respectively)
reveals higher counts of non-local female-sexed bodies of the *Adultus* age category.

<!-- Methods of neighborhood relationship construction examined include
distance based approaches where all the neighbors that fall into a given buffer are considered,
the other two are based on neighborhood graphs, the Delaunay triangulation as a
maximum number of neighbors scenario and a Gabriel graph as a minimal scenario. -->
<!--
For relationships (see Fig. \@ref(fig:nb-sex) and Tab. \@ref(tab:nb-tab-sex))
of male-sexed bodies, none of the methods showed any significant
differences between expected and observed number of neighbors of a given sex.
The relationship with female-sexed bodies is rather extreme for all of the methods employed,
but never significant ($p > 0.07$).
In case of bodies unsexed due to early age (non-adults), the neighborhood based
on buffer zones show a significant relationship with other non-adults ($p \approx 0.04$).
Finally burials with female-sexed bodies show a significantly higher number of
burials with male-sexed bodies in both the methods based on neighborhood networks
(Delaunay triangulation $p \approx 0.02$ and Gabriel graph $p \approx 0.03$).
Two further relationships showed non-random deviations.
In case of Delaunay triangulation for burials with bodies with indeterminate sex with
non-adults ($p \approx 0.02$) and in case of Gabriel graph between non-adults and
indeterminate category $p \approx 0.02$.
-->

<!-- ```{r t-vstat} -->
<!-- vstat <- read_csv(here("data/temp", "v_statistics.csv")) %>% -->
<!--     select(full, v, p) %>% -->
<!--     mutate(v = round(v, 2), p = round(p, 2)) %>% -->
<!--     rename("variable" = full, "v statistic" = v, "p value" = p) -->

<!-- knitr::kable( -->
<!--   vstat, -->
<!--   caption = 'Fuuuuu' -->
<!-- ) -->
<!-- ``` -->

<!-- ======================================================================= -->

# Conclusions and discussion

Our goal was to identify non-random differences between burials both in terms
of the artifact contents of the graves and their position within the cemetery.
The burials or, more precisely, the buried bodies were divided into categories
based on three variables commonly connected to identity in the Neolithic period.
These categories are biological sex, age and origin.
The exploration was conducted both for the categories as single variables and for
all of their combinations.

<!-- Artifact combinations -->
In the case of the grave good analysis, we explored the associations between artifacts and
different categories of the bodies not to establish or reveal any new relationships
but merely to explore whether the known associations between sexes and specific grave goods
are influenced by factors such as age and the local or non-local origin of the buried person.
We did not explore the relationships between grave goods and burial attributes at the
individual artifact scale, so to say; instead, we used an overall measure (Euclidean distance)
of between burial distance in terms of non-randomly distributed artifacts.
By establishing whether the mean ED between categories defined by the grouping attributes
mentioned above falls within or defies the expected distribution derived by
randomization experiment, we identified deviations that are not random.

The most frequently reappearing is the difference between local adult male-sexed bodies and
both local and non-local female-sexed ones and unsexed juveniles.
This relationship shows that sex classification of the body, although being
accentuated by local origin and age category, structures the artifact assemblage
in the grave very well.

<!-- Spatial pattern -->
The spatial organization of the cemetery was explored using point pattern analysis
techniques.
The simpler methods based on cumulative frequency distribution of nearest neighbors
(*G* function) and its generalization for empty space points (*F* function)
did not show deviations from complete spatial random processes.
On the other hand, the methods, including all the distances between events in the
given area  (*K* and *T* functions), indicated that clustering processes are taking place.
The *K* function for multivariate point patterns was also used in order to identify
clustering processes between events of different categories.
We consider this not useful in most cases because the distances at which interesting
clustering processes are suggested imply the existence of a single cluster across
the cemetery and a finer scale is of much greater interest for our work.

Thus, to identify whether close burials are more similar in terms of the categories
outlined above, two neighborhood graphs – Delaunay triangulation and a Gabriel
graph – were used.
They are examples of high and low neighbor case scenario, respectively.
Most of the observations from the graphs proved not significantly different from
the randomization experiment.
There are associations between non-local female-sexed
bodies of the *Maturus* age category and (local) juveniles in both of the graphs.
This observation goes hand in hand with the association of female-sexed bodies
and non-adults in the Delaunay triangulation.
This is not trivial because in general non-adult bodies are
concentrated in the southeast part of the cemetery, whereas female-sexed
bodies are scattered all around the area of the cemetery.
Similarly, individuals determined as belonging to the *Maturus/Senilis* age category are in contrast to
juvenile bodies concentrated in the western half of the cemetery.
Furthermore, the Gabriel graph shows a clear association between non-local *Adultus*
female-sexed bodies and local *Maturus* male-bodies.
This corresponds to the general association of female- and male-sexed bodies of
the *Adultus* and *Maturus/Senilis* age categories in the Gabriel graph.

The most detailed view of the clustering in the cemetery and what kinds of
bodies are prevalent in the given clusters is achieved by the percolation analysis.
The largest cluster across several percolation distances consists of more
non-adult/juvenile bodies and non-local female-sexed bodies of the *Maturus* category
on the one hand while, on the other, the male-sexed bodies are underrepresented here.
This is in full accordance with previous observations.
The central cluster visible at small percolation distances shows the overrepresentation of
local male- and female-sexed bodies.
Also, the southwest cluster has higher counts of local male-sexed bodies of the *Maturus* age category and
a low number of non-local bodies.
On the other hand, non-local female-sexed bodies of the *Adultus* age category are
overrepresented in the western cluster neighboring with the previously mentioned one.

The spatial analysis clearly proved that there is a clustering process taking place
in the cemetery and what more, the age, sex and origin of the buried bodies play
structuring roles in the arrangement of the burials.
The patterns are not too strict, but some appear to be quite constant across the a
range of applied methods.

<!-- ======================================================================= -->

<!-- # Discussion -->

We have successfully shown that all of the categories considered
to be somehow related to identity in the Neolithic and especially their combinations
play substantive roles in structuring the assemblage of grave goods deposited
within the burials.
Some of the relationships between the explored categories and
artifacts are well known at the LBK burial grounds,
for instance, the association between
male burials and polished stone adzes [e.g., @bickle2020; @bentley2012], and some
are known specifically from the *Široká u lesa* cemetery in Vedrovice, e.g.,  the dichotomy
between globular pottery shapes being associated with female-sexed bodies and,
on the other hand, bowls with male-sexed bodies [@kvetina2004].
Our goal was not to establish any new connections between genders or other burial
attributes and grave good categories, but to use these associations
to see which groups of burials are well separated based on the grave goods.

First of all, we have shown that the overall distribution of the artifacts in the graves
was not random; the grave goods distributed non-randomly were identified and
subsequently only these were used in the further analysis, effectively removing any
possible *noise* from artifacts distributed randomly.
Our main premise here is that the identity of the dead is based mostly on attributes like sex,
age and origin defined on the basis of the strontium isotope signal.
We view these *layers* of identity as responsible for differences in
artifact compositions of different graves as well as, to a certain degree, a structuring
factor for the overall spatial organization of the cemetery.
In the case of the grave goods, this relationship is largely dominated by a category
of adult local male-sexed bodies that are distinctly different from both female-sexed bodies
of all age categories and juveniles.
This is caused by the aforementioned artifact characteristic of male burials, the stone adze.
@bickle2020 completely removes polished and beveled stone tools from her analysis
in order to remove their dominance over other patterns.
In our case, adzes are one of four non-randomly occurring artifacts and removing
them from the analysis completely would render it difficult to note any
of the visible patterns.

In the case of spatial analysis, we explore several approaches.
The nearest-neighbor methods are a great tool in analyzing the overall cemetery, but
on a small scale they do not prove useful for discovering any interesting relationships
between studied *layers* of identity.
By far the most useful method in analyzing the intra-site spatial arrangement
is percolation analysis.
Another important point is the use of resampling methods that allow for the identification
of random and non-random patterns, both in grave good analysis and spatial organization.

<!--Traditional GIS based methods were already applied to study the cemetery
[@kvetina2004] the techniques applied here reveal some trends that were previously-->

We have shown that not only sex, which is traditionally perceived
as the most important structuring proxy for identity differentiation, but also
age and, quite importantly, also the origin of the deceased played substantial
roles in equipping the graves with various types of artifacts by the Vedrovice community.
In regard to the spatial arrangement of the cemetery, we demonstrated that there are
prevalent clustering processes taking place across the cemetery and also groups
based on certain identities tend to be buried in clusters.
These observations are based on the extensive use of resampling approaches that
allowed us to distinguish between patterns occurring solely by chance and
patterns significantly deviating from randomness.

<!-- ======================================================================= -->
<!-- FIGURES --><!-- FIGURES --><!-- FIGURES --><!-- FIGURES --><!-- FIGURES -->

```{r fig-widths}
fig_wid <- c(".3\\textwidth",
             ".6\\textwidth",
             ".8\\textwidth",
             "\\textwidth")
```

<!-- randomization -->
```{r sstat, fig.cap="Distribution of simulated and observed *S* statistic", out.width=fig_wid[1]}
knitr::include_graphics(here("plots/cooc_sstat.pdf"))
```

```{r cooc-vstat, fig.cap="Simulated and observed *V* statistics for different variables", out.width=fig_wid[4]}
knitr::include_graphics(here("plots", "cooc_vstats.pdf"))
```

```{r label-cooc-nets}
label_cooc_nets <- paste("Networks of variables frequently present together and excluding each other.",
                         "1 – bottle, 2 – bowl, 3 – globular pot, 4 – polished stone tool, 5 – local lithics,",
                         "6 – non-local lithics, 7 – pebble, 8 – grinding slab, 9 – bone tool,",
                         "10 – U-shaped pendant, 11 – spondylus beads, 12 – O-shaped buckle.",
                         "Images adapted from Podborský 2002")
```
```{r cooc-nets, fig.cap=label_cooc_nets, out.width=fig_wid[4]}
knitr::include_graphics(here::here("plots", "cooc_nets_fin.jpeg"))
```

<!-- plans -->
```{r plan-sex, fig.cap="Plan of the cemetery, kernel density estimation for differently sexed bodies", out.width=fig_wid[4]}
knitr::include_graphics(here("text/figs/plan_sex.pdf"))
```

```{r plan-age, fig.cap="Plan of the cemetery, kernel density estimation for bodies of different ages", out.width=fig_wid[4]}
knitr::include_graphics(here("text/figs/plan_age.pdf"))
```

```{r plan-orig, fig.cap="Plan of the cemetery, kernel density estimation for bodies of different origins", out.width=fig_wid[2]}
knitr::include_graphics(here("text/figs/plan_origin.pdf"))
```

<!-- point pattern analysis -->

```{r label-pp-funs}
label_pp_funs <- paste(
  "Functions *G*, *F*, *K* and *T*.",
  "*Dashed line* -- theoretical curve under complete spatial randomness,
  *solid line* -- empirical curve, *gray area* -- envelope")
```
```{r pp-funs, fig.cap=label_pp_funs, out.width=fig_wid[4]}
knitr::include_graphics(here("text/figs/pp_funs.pdf"))
```

```{r label-pp-crossk}
label_crossk <- paste("Multitype K function for different categorical variables; for legend, see previous Figure")
```
```{r pp-crossk, fig.cap=label_crossk, out.width=fig_wid[3]}
knitr::include_graphics(here("text/figs/pp_crossk.pdf"))
```

<!-- neighborhood -->
```{r plan-graphs, fig.cap="Neighborhood graphs: Delaunay triangulation and Gabriel graph", out.width=fig_wid[4]}
knitr::include_graphics(here("text/figs/plan_graphs.pdf"))
```

<!-- percolation analysis -->

```{r label-perc-dist}
label_perc_dist <- "Normalized sizes of largest clusters at given distances, percolation thresholds labeled"
```
```{r perc-distance, fig.cap=label_perc_dist, out.width=fig_wid[2]}
knitr::include_graphics(here("text/figs/perc_distance.pdf"))
```

```{r perc-plan, fig.cap="Clusters formed at given percolation threshold distances", out.width=fig_wid[3]}
knitr::include_graphics(here("text/figs/perc_plan_ids.pdf"))
```

<!-- ======================================================================= -->

\pagebreak

# References